Stack Safety as a Security Property

### 2020-07-08 main.tex
CP for rob

\paragraph*{Testing the conservative policy}
%
This conservative policy can be shown to satisfy an eager \emph{testing
property}, which at each step of the computation performs integrity and
confidentiality tests for a stack of running variants, which are kept in
lockstep with the original machine.

\definition{Stack integrity test}
  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
  satisfy the \emph{stack integrity test} if their high-integrity components
  match:
  \[
    \mathit{int}_T ~ \contour ~ \mach ~ \mach' \triangleq
    \forall \component . ~
    \contour(\component) = \HIGHINT \rightarrow
    \mach(\component) = \mach'(\component)
  \]

\definition{Stack confidentiality test}
  Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
  and observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
  \stepstoobs{\obs_2} \mach_2'$, all these satisfy the \emph{stack
    confidentiality test} if their observations and changed components coincide:
  \[
    \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2' ~ \triangleq
    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1'(\component) = \mach_2'(\component)
  \]
\apt{scoping of $\forall$ is ambiguous.}
\rb{Fix notations, including for properties below.
  In particular, note that the last four arguments could be computed from the
  first two and it may be clearer to elide them.}
\newcommand{\vse}{\mathit{vse}}
\newcommand{\VSES}{\mathit{VSE}}
\newcommand{\vs}{\mathit{vs}}
\newcommand{\VSS}{\mathit{VS}}

\definition{Variant stacks}
  The testing property keeps track of relevant information on a stack whose
  elements represent variants of the machine by recording an initial machine
  state, an initial and current variant of the machine, a contour and a return
  map:
%
  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
  $$\vs \in \VSS$$
%
  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
  respect to it if the stack is nonempty, the return map of the bottom element
  is the false property, and for each element $\vse \in \vs$:
  \begin{itemize}
  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
  \item $\mach$ is in the trace $\machT$ of its initial machine,
    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
  \item Its variant state $\mathit{curr\_variant}(\vse)$
    is in the trace of its initial variant,
    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
  \item The return function is in the return map if not the last element. \rb{TODO}
  \item The registers of its contour $\contour(\vse)$ are public.
  \end{itemize}

\definition{Variant stack test}
%
  Our testing property extends the basic tests above to whole variant stacks.
  Given a machine that steps $(\mach, \pol) \hookrightarrow_P (\mach', \pol',
  \obs)$, say that a variant stack $\vs$ that is well-formed with respect to
  $\mach$ safisfies the \emph{variant stack test} if for each variant element
  $\vse$, the element can take a step $\mathit{curr\_variant}(\vse)
  \stepstoobs{\obs} \mach'$ while satisfying the stack integrity test
  $\mathit{int}_T ~ \contour(\vse) ~ \mach ~ \mach'$ and stack confidentiality
  test $\mathit{conf}_T ~ \mach ~ \mach'$ with regard to the original machine
  state. We write this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.

>>>>

\paragraph*{Strong Stack Safety}
%
This conservative policy does not only satisfy the stack safety above,
but also a provably stronger one, which at each step of the
computation performs integrity and confidentiality tests for a stack
of running variants, which are kept in lockstep with the original
machine.

\definition
  Given two machine states $\mach$ and $\mach'$ and a contour $\contour$, these
  satisfy the \emph{stack integrity test} if their high-integrity components
  match:
  \[
    \mathit{int}_T ~ \contour ~ \mach ~ \mach' \triangleq
    \forall \component . ~
    \contour(\component) = \HIGHINT \rightarrow
    \mach(\component) = \mach'(\component)
  \]

\definition
  Given two machine states $\mach_1$ and $\mach_2$ with stepping machine states
  and observations $\mach_1 \stepstoobs{\obs_1} \mach_1'$ and $\mach_2
  \stepstoobs{\obs_2} \mach_2'$, all these satisfy the \emph{stack
    confidentiality test} if their observations and changed components coincide:
  \[
    \mathit{conf}_T ~ \mach_1 ~ \mach_2 ~ \obs_1 ~ \obs_2 ~ \mach_1' ~ \mach_2' ~ \triangleq
    \obs_1 = \obs_2 \wedge \forall \component . ~
   (     \mach_1(\component) \neq \mach_1'(\component)
    \vee \mach_2(\component) \neq \mach_2'(\component))
    \rightarrow \mach_1'(\component) = \mach_2'(\component)
  \]
\apt{scoping of $\forall$ is ambiguous.}
\rb{Fix notations, including for properties below.
  In particular, note that the last four arguments could be computed from the
  first two and it may be clearer to elide them.}
\newcommand{\vse}{\mathit{vse}}
\newcommand{\VSES}{\mathit{VSE}}
\newcommand{\vs}{\mathit{vs}}
\newcommand{\VSS}{\mathit{VS}}

\newcommand*{\vm}[1]{\ensuremath{\mathit{vm}_{#1}}}

\definition The testing property keeps track of relevant information
on a {\em variant stack}: a list of machine states, each element of
which (denoted $\vm{C}$) of which is a variation of the actual machine
according to a contour $C$.
%
\footnote{The full formal definition in Coq also records additional
  information (the initial machine state at the call, its initial
  variation, the contour, and a return predicate) which we elide
  here for simplicity of exposition.}
%%
%  $$\vse \in \VSES ::= \MACHS \times \MACHS \times \MACHS \times \CONTOURS \times \RETMAPS$$
%  $$\vse = (\mathit{init\_machine}(\vse), \mathit{init\_variant}(\vse), \mathit{curr\_variant}(\vse), \contour(\vse), \mathit{ret}(\vse))$$
%  $$\vs \in \VSS$$
%%
\iflater
\leo{This can be elided}
  Given a machine state $\mach$, a stack $\vs$ is \emph{well-formed} with
  respect to it if the stack is nonempty, the return map of the bottom element
  is the false property, and for each element $\vse \in \vs$:
  \begin{itemize}
  \item $\mathit{curr\_variant}(\vse)$ is a variant of $\mach$.
  \item $\mach$ is in the trace $\machT$ of its initial machine,
    $\mathit{init\_machine}(\vse) \hookrightarrow \machT$.
  \item Its variant state $\mathit{curr\_variant}(\vse)$
    is in the trace of its initial variant,
    $\mathit{init\_variant}(\vse) \hookrightarrow \machT$.
  \item The return function is in the return map if not the last element. \rb{TODO}
  \item The registers of its contour $\contour(\vse)$ are public.
  \end{itemize}
  \fi

\definition
%
  Our testing property extends the basic tests above to whole variant
  stacks.  Given a machine that steps $(\mach, \pol) \hookrightarrow_P
  (\mach', \pol', \obs)$, say that a variant stack $\vs$ \iflater that
  is well-formed with respect to $\mach$\fi safisfies the
  \emph{variant stack test} if for each variant element $\vm{C}$
  (which is a variation of $\mach$ for $C$), $\vm{C}$ can take a step
  $\mathit{curr\_variant}(\vse) \stepstoobs{\obs} \mach'$ while
  satisfying the stack integrity test $\mathit{int}_T ~ \contour(\vse)
  ~ \mach ~ \mach'$ and stack confidentiality test $\mathit{conf}_T ~
  \mach ~ \mach'$ with regard to the original machine state. We write
  this $\mathit{\vs}_T ~ (\mach, \pol) ~ \vs$.
